---
title: "R/C++"
output:
 html_document:
  toc: true
  toc_depth: 2
  toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(microbenchmark)
```
I mostly code in __R__, __R/C++__, __Python__ and __VBA__ depending upon the situation, purpose and need. I am  *language agnostic* (although I will vigorously avoid having to read/write Perl, if I can). At various times I have had reason to use __Java__ (mostly for agent-based modelling via the [MASON](http://cs.gmu.edu/~eclab/projects/mason/) framework), __SQL__ for data gathering and manipulation, and __Perl__ for basic scripting. 

For specialised mathematical tasks I have used __Matlab__ and __Mathematica__ (and the intriguing [Wolfram language](http://www.wolfram.com/language/fast-introduction-for-programmers/en/)). I experimented with [Julia](http://julialang.org/), in its (very) early days. 

Languages I find interesting (for one or another reasons) but have not (yet) needed to use, are:

* [__D__](https://dlang.org/) -- a language that started out as re-engineered C++, that claims to be safer and easier to code than its forebear.

* [__Clojure__](https://clojure.org/) -- a dialect of Lisp, a language as old as Fortran, but still very much in use today. Supposedly supports functional programming and designed for concurrency.

* [__Ruby__](https://www.ruby-lang.org/en/) -- an interesting and well-regarded general purpose language developed in the 1990s.

* [__C#__](https://msdn.microsoft.com/en-gb/library/67ef8sbd.aspx) -- a well-known and popular language similar to C++ and Java.

* [__Haskell__](https://www.haskell.org/) -- probably the most *heard-of* language in this list. A purely functional language named after Haskell Curry, who worked in combinatorial logic. Even the description sounds *difficult*...

# R

I have used R since 2003, where I picked it up for statistical computing (and applied statistics). While the core language has not changed too much since then, the R ecosystem (meaning: developer environments, language interoperability and extensions, packages and domain specific use-cases) has grown rapidly over the last few years.

R is considered to be an *unconventional* programming language. While most beginners often think of it as a fancy statistical calculator, it is both a powerful and flexible language / scripting language for the more general R environment (similar to how VBA is a scripting language for Excel). Although this definition downplays many of R strengths - it is accurate. For casual R users, much of the advanced mechanisms are not immediately obvious: many new R users learn how to write and use (user-defined) functions quite quickly, without having to know [these details](https://kevinushey.github.io/blog/2015/11/22/what-is-a-function/), for instance.

* There are many guides to R on-line, and much of the basic syntax and functionality can be picked up using a combination of *doing*, reading the source code and (excellent) documentation of packages. A great book for newcomers to R is [Learning R](https://www.amazon.co.uk/Learning-R-Richard-Cotton/dp/1449357105/ref=sr_1_2?s=books&ie=UTF8&qid=1488064624&sr=1-2) by Richard Cotton. For intermediate users, or those interested in the more advanced features of the language, Hadley Wickham's [Advanced R](http://adv-r.had.co.nz/) is hard to beat. 

## Performance: Integrated R & C++ via Rcpp

[Rcpp](http://dirk.eddelbuettel.com/code/rcpp.html) is a R package enabling the *easy* use of compiled code within R. While the R environment makes extensive use of compiled C and Fortran, including compiled code in packages and user-defined functions was, until now, *painful*. It is difficult to overstate how workflow in R is extended by Rcpp. Along with the optimised matrix operations of the [Armadillo C++ library](http://arma.sourceforge.net/), available to R through the [RcppArmadillo](http://dirk.eddelbuettel.com/code/rcpp.armadillo.html) package, and the recent [RcppParallel](https://rcppcore.github.io/RcppParallel/) package, leveraging CPU multi-core concurrency direct from compiled C++, R is a serious environment for numerical programming. A good introduction to RcppParallel is [this talk](https://www.rstudio.com/resources/webinars/rcppparallel-a-toolkit-for-portable-high-performance-algorithms/).

A trivial example of how simple it is to deploy solutions in C++ 
```{r engine='Rcpp',eval=FALSE}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::plugins("cpp11")]]

// [[Rcpp::export]]
int i_sect(const IntegerVector& x, const IntegerVector& y ) {
  std::unordered_set<int> s(x.begin(), x.end());
  return std::count_if(y.begin(), y.end(), [&](int k) {return s.find(k) != s.end();});
}

// [[Rcpp::export]]
double i_prod(NumericVector& x, NumericVector& y) {
  return std::inner_product(x.begin(), x.end(), y.begin(),0.0);
}
```


```{r output_rcp,eval=FALSE}
require(microbenchmark)
x<-runif(100000)
microbenchmark(sum(x * x),i_prod(x,x))
```
Resulting in a 2x to 4x speed gain. While C++ provides performance benefits, making previously infeasible numerically intensive simulation tasks feasible, it also exposes data structures and algorithms to R.

<br>

Here is a simple template function to sort a vector relative to another. This is 2x as fast as the naive R method, and even faster for small vectors.
```{r sort, engine='Rcpp', ,eval=FALSE}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::plugins("cpp11")]]

template <typename T, typename U> void sorty(T& a, U const x) {
  std::sort(a.begin(), a.end(),
            [&x](int i, int j) -> bool { return x[i] > x[j]; });
}

// [[Rcpp::export]]
NumericVector sort_by(NumericVector& to_sort, NumericVector& by){
  IntegerVector toOrder = seq_len( by.size() )-1; //similar to std::iota, gives {0,1,2,..,by.size()-1}
  sorty(toOrder,by);
  return(to_sort[toOrder]);
}
```

```{r output_rcp2,eval=FALSE}
a<-rexp(1000,1)
b<-rnorm(1000,0,1)
microbenchmark(a[order(b,decreasing = TRUE)],sort_by(a,b))
```
<br>

## Deployment: Shiny and Knitr

[Shiny](https://shiny.rstudio.com/) is a web application framework enabling the deployment of interactive R sessions over the web. [KnitR](https://yihui.name/knitr/) is a dynamic report generator for R, allowing documents (html, latex, pdf, Word documents) to be built around users workflow in R and dynamically capture results from R computations. Combining these together with [R Markdown](http://rmarkdown.rstudio.com/index.html) documents and RStudio notebooks, makes dissemination and deployment capabilities of R exceptionally strong.

### Stick breaking with Shiny

A Shiny interactive app to investigate the expected lengths of each of ordered $n+1$ segments of unit length stick. Tap the picture below to be taken to the Shiny app, or follow this [link](https://chris-x.shinyapps.io/StickBreak/).

<!--- this actually works---!>
<!-- <iframe src="https://chris-x.shinyapps.io/StickBreak/" style="border: none; width: 440px; height: 600px"></iframe> -->


[![[Shiny app](https://chris-x.shinyapps.io/StickBreak/)](images/stick_io.jpg)](https://chris-x.shinyapps.io/StickBreak/)

## Data: tidyverse and data.table

[Tidyverse](https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/) is the term given to the suite of R packages that have significantly enhanced (to put it mildly) how one approaches the task of data analysis, data wrangling and data visualisation in R -- compared to the *base R* workflow of `data.frame` and `plot`. While data frames are still a core feature of the language, and base R plot can plot almost anything (if you dedicate enough time to the task), the tidyverse philosophy significantly reduces the time (and effort) it takes to produce those *interesting* and *complex* analyses base R struggles with.

[Data.table](https://github.com/Rdatatable/data.table/wiki) offers another replacement for base data frames, utilising *by-reference* operations on data, it is incredibly fast and efficient and can work with GB's of data effortlessly. While dplyr (part of the tidyverse suite) offers a clean and consistent syntax - together with much improved performance over direct use of `data.frame` - data table should be considered where speed is important.

<br>

## Machine learning
