---
title: "Research"
output:
 html_document:
  toc: true
  toc_depth: 2
  toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(visNetwork)
library(igraph)
```
```{r networks, echo=FALSE}
colfunc<-colorRampPalette(c("springgreen","purple4"))

iggB<-barabasi.game(60,1,directed=FALSE)
V(iggB)$name=as.character((1:vcount(iggB)))

iggA<-sample_grg(20,0.28,torus = TRUE)
V(iggA)$name<-as.character(1:vcount(iggA))

igg<-union(iggB, iggA)
GEdge<-get.edgelist(igg)
GEdgedf<-data.frame(from=GEdge[,1],to=GEdge[,2],weight=rep(1,nrow(GEdge)),color=rep("purple",nrow(GEdge)))

DD<-degree(igg);
Gnodes <- data.frame(id = as.numeric(V(igg)$name),value=DD,color=colfunc(10)[cut(DD,breaks=10,labels=FALSE)])
visNetwork(Gnodes,GEdgedf,main = "scale-free network with geometric core") %>% visInteraction(dragView=FALSE,zoomView=FALSE)
```

# Complex systems and networks

I am primarily interested in complex systems and spreading processes on networks. As a naturally interdisciplinary field of research, tools I have worked with include
<style>
  .col2 {
    columns: 2 150px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 150px; /* chrome, safari */
    -moz-columns: 2 150px;    /* firefox */
  }
</style>

<br>

<div class="col2">

- Graph and network theory
- Stochastic and probabilistic analysis
- Dynamical systems
- Combinatorics

- Simulation and computation
- Agent-based models
- Statistical physics
- Percolation
</div>

<br>

I am equally interested in applying my research to financial markets and in economic contexts.

<br>

# Selected Publications & reports

- [Cascades on a stochastic pulse-coupled network](http://www.nature.com/articles/srep06355). Wray, C., Bishop, S. 2014

- A mathematical framework for agent-based models. Wray, C., Knipl, D., Bishop, S. (2015 Working paper)

- [A Financial Market Model Incorporating Herd Behaviour](http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0151790). Wray, C., Bishop, S. 2016

- Kinetics of social contagion on temporal interconnected networks. Wray, C., Knipl, D., Bishop, S., Kertész, J. (2017 Working paper)




